---
title: "Giant Pacific Salamanders"
author: "Waldo and Wes"
date: "12/2/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

```
Introduction: 

Summary: 

Photos: 



```{r}
#Attach Packages
library(tidyverse)
library(janitor)
library(kableExtra)
library(effsize)
```



```{r}
#Read in Salamander Data
MC_Salamanders <- read_csv("mack_creek_vertebrates.csv") %>% 
    clean_names() %>% 
  filter(species == "DITE") %>%  
           mutate(river_type = (case_when(
            unittype == "C" ~ "Cascade", 
            unittype == "SC" ~ "Side Channel",
            unittype == "P" ~ "Pool"))) %>% 
  mutate(forest_type = (case_when(
            section == "OG" ~ "Old Growth",
            section == "CC" ~ "Clear Cut")))
```


```{r}
#Wrangle Data (Result A)
MC_annual_count <- MC_Salamanders %>%  
  filter(species == "DITE") %>% 
  group_by(year,forest_type) %>% 
  summarize(annual_count = n())
```


```{r}
#Graph Data (Result A)
ggplot(MC_annual_count, aes(x = year, y = annual_count)) +
geom_line(aes(color = forest_type),
            size = .5,
            show.legend = TRUE) + 
  geom_point(aes(year))+
   scale_x_continuous(limits = c(1993, 2017),
                     breaks = seq(1993, 2017, by = 1)) +
  scale_y_continuous(limits = c(0, 400),
                     breaks = seq(0, 500, by = 50)) + 
  theme_minimal()+ 
  labs(title = "Pacific Giant Salamander Counts in Old Growth and Clear Cut Forest",
       subtitle = "Mack Creek, HJ Andrews Experimental Forest, OR",
       x = "Year",
       y = "Salamander Count", 
       color = "Forest Condition") +
  theme(axis.text.x = element_text(angle= 45, 
                                   hjust = 1),
        axis.ticks = element_line(size = .5)) 
  
```
**Figure 1** *Salamander counts were recorded in both clear-cut and old-growth forests abutting Mack Creek for 25 years.*



```{r}
#Wrangle Data (Result B)
MC_2017_count <- MC_Salamanders %>% 
  filter(year == "2017") %>% 
  filter(unittype %in% c("P", "C", "SC")) %>% 
  count(river_type, forest_type)

MC_table <- MC_2017_count %>% 
  pivot_wider(names_from = river_type, values_from = n)

MC_proportion <- MC_table %>% 
  adorn_percentages(denominator = "row") %>% 
  adorn_pct_formatting(digits = 0)  %>% 
  rename("Forest Condition" = forest_type)
```



```{r}
#Table Data (Result B)
kable(MC_proportion) %>% 
  kable_styling(bootstrap_options = "basic") 
```
**Figure 2.** *Proportion of salamanders found in different stream types within clear-cut and old-growth forests during 2017.* 


```{r}
#Wrangle Data (Result C) 
Sal_chi_test <- MC_table %>%  
  pivot_longer(cols = c("Cascade","Pool", "Side Channel")) %>% 
  pivot_wider(names_from = forest_type) %>% 
  select(-name)
```



```{r}
#Chi-Squared Test (Result C)
Sal_chi_results <- chisq.test(Sal_chi_test)
```
Forest conditions do not affect the proportion of salamanders found in cascades, pools, or side channels. 
($\chi$^2^(`r Sal_chi_results$parameter`) = `r round(Sal_chi_results$statistic,2)`, p = `r round(Sal_chi_results$p.value, 3)`).
